```{r}
#Load Data
library(ISLR)
set.seed(1)
data("Wage")
```

Run linear regression on training set.
```{r}
train <- sample(dim(Wage)[1], dim(Wage)[1]/2)
cv_linear <- lm(wage~as.factor(education)+poly(age,1), data = Wage, subset = train)
yhat_error_linear <- Wage$wage-predict(cv_linear, newdata = Wage)
MSE_linear <- mean(yhat_error_linear[-train]^2)
MSE_linear
```

Run quadratic regression on training set.
```{r}
train <- sample(dim(Wage)[1], dim(Wage)[1]/2)
cv_quadratic <- lm(wage~as.factor(education)+poly(age,2), data = Wage, subset = train)
yhat_error_quadratic <- Wage$wage-predict(cv_quadratic, newdata = Wage)
MSE_quadratic <- mean(yhat_error_quadratic[-train]^2)
MSE_quadratic
```

Run cubic regression on training set.
```{r}
train <- sample(dim(Wage)[1], dim(Wage)[1]/2)
cv_cubic <- lm(wage~as.factor(education)+poly(age,1), data = Wage, subset = train)
yhat_error_cubic <- Wage$wage-predict(cv_cubic, newdata = Wage)
MSE_cubic <- mean(yhat_error_cubic[-train]^2)
MSE_cubic
```

Generate k-folds
```{r}
set.seed(1)
K <- 2
fold <- sample(1:K,length(Wage$wage), replace = T)
error <- Wage$wage
```

Run regression excluding fold 1
```{r}
wage_lm <- lm(wage~poly(age,1)+as.factor(education),data=Wage, subset = (fold != 1))
wage_hat <- predict(wage_lm, newdata = Wage)
error[fold == 1] <- (Wage$wage - wage_hat)[fold == 1]
```

Run regression excluding fold 2
```{r}
wage_lm <- lm(wage~poly(age,1)+as.factor(education),data=Wage, subset = (fold != 2))
wage_hat <- predict(wage_lm, newdata = Wage)
error[fold == 2] <- (Wage$wage - wage_hat)[fold == 2]
```

Calculate cross-validated MSE
```{r}
(1/K) * (mean(error[fold == 1]^2) + mean(error[fold == 2]^2))
```

Run regression excluding fold 1
```{r}
wage_quad_lm <- lm(wage~poly(age,2)+as.factor(education),data=Wage, subset = (fold != 1))
wage_quad_hat <- predict(wage_quad_lm, newdata = Wage)
error[fold == 1] <- (Wage$wage - wage_quad_hat)[fold == 1]
```

Run regression excluding fold 2
```{r}
wage_quad_lm <- lm(wage~poly(age,2)+as.factor(education),data=Wage, subset = (fold != 2))
wage_quad_hat <- predict(wage_quad_lm, newdata = Wage)
error[fold == 2] <- (Wage$wage - wage_quad_hat)[fold == 2]
```

Calculate cross-validated MSE
```{r}
(1/K) * (mean(error[fold == 1]^2) + mean(error[fold == 2]^2))
```


Run regression excluding fold 1
```{r}
wage_cubic_lm <- lm(wage~poly(age,3)+as.factor(education),data=Wage, subset = (fold != 1))
wage_cubic_hat <- predict(wage_cubic_lm, newdata = Wage)
error[fold == 1] <- (Wage$wage - wage_cubic_hat)[fold == 1]
```

Run regression excluding fold 2
```{r}
wage_cubic_lm <- lm(wage~poly(age,3)+as.factor(education),data=Wage, subset = (fold != 2))
wage_cubic_hat <- predict(wage_cubic_lm, newdata = Wage)
error[fold == 2] <- (Wage$wage - wage_cubic_hat)[fold == 2]
```

Calculate cross-validated MSE
```{r}
(1/K) * (mean(error[fold == 1]^2) + mean(error[fold == 2]^2))
```

```{r}
set.seed(1)
library(boot)
for (i in 1:4) {
  wage_glm <- glm(wage~poly(age,i)+as.factor(education), data=Wage)
  cv.err <- cv.glm(data=Wage,glmfit = wage_glm, K = 5)
  print(cv.err$delta[1])
}
```





